---
title: "General Social Survey"
author: "Melanie Vera"
date: "September 10, 2024"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
source("GrabData.R")
```




#  Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data in the dataframe GSSdata is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 16 variables. It has further removed unwanted factor levels in much of the data. Examine the code in the GrabData.R file to see what it is doing. Some of the variables are categorical and others are numerical. Be sure to do a variable analysis before tackling each question.  
First question - Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
Question: is gun ownership independent of opinion on capital punishment?

$H_0$: gun ownership and opinion on capital punishment are not related.

$H_A$: gun ownership and opinion on capital punishment are related.


OWNGUN ~ CAPPUN both are categorical, thus is use cat~cat


## Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~OWNGUN + CAPPUN,data=GSSdata)
barchartGC(~OWNGUN + CAPPUN,data=GSSdata, type="percent")

```


This bar chart shows gun owners are more likely to favor capital punishment.The percent being over 55% for people that voted yes and no.


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
table1 <- xtabs(~OWNGUN + CAPPUN, data=GSSdata)
rowPerc(table1)
colPerc(table1)
```

75.68 percent of people are in favor.
### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

We got the data from the cat~cat file. The null hypothesis was rejected due to the p-value=0.0202. Meaning that people were more likely to favor capital punishment, 75.68. Gun ownership and capital opinion are related.

#  Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?

$H_0$: There is no difference in years of education between those who did and didn't vote in 2012

$H_A$: There is a difference in years of education between those who did and didn't vote in 2012.



## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
boxplot(EDUC~VOTE12,data=GSSdata)
```


The box plot shows that people that voted in 2012 had an average of 16 years of education and people that didn't vote had an average of 13 years.


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(EDUC~VOTE12,data=GSSdata)
```
There is a mean of 13 for both people that voted in 2012 and people that didn't vote.People who voted in 2012 on average had 16 years of education.

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfy the assumptions and is appropriate to answer the research question!-->

```{r}
t.test(EDUC~VOTE12,data=GSSdata)
```

We reject the null hypothesis. There is a difference in years of education between those who did and didn't vote in 2012 because there was 15.72656 people that voted and 13.42000 people that didn't vote.

#  Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).

$H_0$: A persons performance on the WORDSUM are related to their level of schooling DEGREE.

$H_A$: A persons performance on the WORDSUM are not related to their level of schooling DEGREE.

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
plot(WORDSUM~DEGREE,data=GSSdata)
```


The box plots show a slight difference in data for each group. LT high school has a mean of 7, Junior college has a mean of 9, and and graduate has a mean of 10. Meaning that people in LT High school scored the lowest because of there level of schooling degree.


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=GSSdata)
```

The table reflects that Junior college and bachelor are approximately the same mean. Where the graduates is much higher and the LT High school and high school is significantly lower.

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
model1 <- aov(WORDSUM~DEGREE,data=GSSdata)
summary.aov(model1)
```
We reject the null hypothesis since he p-value is lower than 0.5. There is a difference between the average score on the WORDSUM  and the EDUC. 

```{r}
plot(TukeyHSD(model1))
```

The turkey test shows us the differences in mean levels of DEGREE between graduate bachelor and Bachelor LT High School. Showing that 5 lines are touching the 0.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variables causes the other, or whether they perhaps are related for some other reason.  Watch the chapter 6 videos from the George Town videos collection.!-->

Q1: In conclusion the null hypothesis was rejected due to the p-value=0.0202. Meaning that people were more likely to favor capital punishment, 75.68. Gun ownership and capital opinion are related.

Q2:In conclusion we rejected the null hypothesis. There is a difference in years of education between those who did and didn't vote in 2012 because there was 15.72656 people that voted and 13.42000 people that didn't vote.

Q3: In conclusion we reject the null hypothesis since he p-value is lower than 0.5. There is a difference between the average score on the WORDSUM  and the EDUC. Junior college and bachelor are approximately the same mean. Where the graduates is much higher and the LT High school and high school is significantly lower.
